### 2015-11-11 paper.tex
[minor] formatting

\documentclass{sig-alternate}

\usepackage{hyperref}

\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{booktabs}

\usepackage[draft,nomargin,footnote]{fixme}

\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning, calc}

\usepackage{xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
\newcommand{\etal}{\textit{et al.}\xspace}
\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\etc}{\textit{etc.}\xspace}
\newcommand{\vs}{\textit{vs.}\xspace}
\newcommand{\redacted}{\textit{\sc [REDACTED FOR REVIEW]}\xspace}

\graphicspath{{figs/}}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{HRI}{'16 Chrischurch, New Zealand.}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction}

%\author{Séverin Lemaignan \qquad Fernando Garcia \qquad Alexis Jacq \qquad Pierre Dillenbourg\\Computer-Human Interaction in Learning and Instruction Laboratory (CHILI)\\École Polytechnique Fédérale de Lausanne (EPFL)}
\author{Redacted for double-blind review}

\maketitle
\begin{abstract}

Measuring ``how much the human is in the interaction'' -- the \emph{level of
engagement} -- is instrumental in building effective interactive robots.

Engagement, however, is a complex, multi-faceted cognitive mechanism that is
only indirectly observable. This article formalizes \emph{with-me-ness} as one
of such indirect measures. \emph{With-me-ness}, a concept borrowed from the
field of {\it Computer-Supported Collaborative Learning}, measures in a
well-defined way to what extent the human is \emph{with} the robot over the
course of an interactive task. As such, it is a meaningful precursor of
engagement.

We expose the full methodology, from real-time estimation of the human's focus
of attention (relying on a novel, open-source, vision-based head pose
estimator), to on-line computation of with-me-ness.

We report as well on the experimental validation of this approach, using a
naturalistic setup involving children during a complex robot-teaching task.

\end{abstract}

%% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Experimentation, Human Factors.}

\keywords{Human-Robot Interaction; Social Robotics; Attention Assessment; Engagement; Computer Vision.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

%In the recent years, research is pushing forward in building intelligent systems
%and environments that try to maximize the benefits of cooperation with humans
%based on verbal and non-verbal communication. The interaction between social
%robots and humans entails a broad set of situations, and evidences the need for
%capturing significant backchannel information generated by the user during the
%interaction process. Such information is used to adapt the system to the
%specific user's needs to eventually provide a social or emotional intelligence
%to the agents.
%
%It is indeed not enough to exhibit human-like capabilities in social tasks: a
%robot needs to achieve human-like manners. For instance, several studies showed
%that humans prefer to interact with robots whose task outcome is delayed, shows
%uncertainty or an undesired output \cite{Admoni,Short}. In fact, the ability to
%adapt has been proved to be beneficial in long-term interactions
%\cite{Tielman:2014, Lim:2014}, specifically with children, manifesting a more
%positive attitude over time.

Building capable social agents requires to endow them with a range of perceptual
capabilities: however, while face and object tracking and recognition, path
planning, speech recognition or task learning are some of the current major
research directions, existing literature tends to evaluate each of these
algorithms in their own metric space, without considering the interaction
quality at a global level. Anzalone~\etal~\cite{anzalone} have recently argued
in favor of such a global evaluation, and they propose to assess these
algorithms in terms of their capability to \emph{obtain the desired effect} in a
human-robot interaction context. They correspondingly propose metrics built
around the measurement of \emph{engagement} as indicator of the quality of the
experience.

``Engagement'', the cognitive, affective and behavioral state of interaction
with a computer application that ``makes the user want to be
there"~\cite{OBrien:2010}, has actively been studied in a diverse set of
domains. Specifically in robotics, several variables and social signals have
been proposed in the literature to quantify it. A recent review of these is
presented in~\cite{ivaldi2015towards}.

%In a learning context, the evaluation of response times in relation with
%correctness has been proposed~\cite{Beck} to provide estimate the user's
%engagement, with however large inter-subject and environment variability.
%Biometric measurements have also been proposed, like the use the weight
%distribution of the user on a pressure-sensitive
%chair~\cite{Chipman07postureas}.

For instance, \cite{Castellano:2009} proposes to predict children's level of
engagement by integrating non-verbal cues (gaze and smiles) with the current
state of the interaction in a Bayesian model. While they report a high level of
accuracy, their approach requires post-hoc video annotations, and is not
applicable to on-line engagement assessment. Similarly,
Baxter~\etal~\cite{baxter2014tracking} posit that the measure of the direction
and timing of gaze in child-robot interactions is a proxy for engagement and
attribution of social agency. However, they also conduct these measures as
post-hoc analyses.  \cite{peters2010investigating} model the user's interest and
engagement with a virtual agent by tracking eye gaze and head direction.
Similarly, \cite{ishii2011combining} estimates the user's engagement with a
conversational agent based on the analysis of gaze patterns.
In~\cite{Rich:2010}, a computational model based on the recognition of
\emph{connection} events such as directed gaze, mutual facial gaze is proposed.
Not relying on gaze, \cite{Sanghvi:2011} focus only on the back or trunk posture
as a determining factor for the assessment. Finally, a recent study with social
robots in face-to-face scenarios~\cite{anzalone} explores a set of metrics based
on non-verbal cues but they also underline possible limitations in long-term
scenarios.

The variety of these approaches reflects the fact that \emph{engagement} remains
a broad concept, fairly ill-defined and thus difficult to operationalize.
Therefore, instead of introducing ``yet another metric of engagement'', we
introduce the more specific concept of \emph{with-me-ness}~\cite{sharma2014me}:
to what extent the human is ``with me'', the robot, during the interaction.

We measure \emph{with-me-ness} by comparing the attentional focus of the human
(as estimated in real-time by the robot) with the expected, {\it a priori}
targets of attention elicited by the task at hand.

The following sections expose the full methodology, starting with the estimation
of attention: we present in the next section a novel method for on-line estimation
of the focus of attention based on fast 6D head pose estimation. We validate
this technique in section~\ref{sec:expe} with a real-world field study involving
children. Section~\ref{sec:withmeness} formally introduces the concept of
\emph{with-me-ness}. We present how to compute it over the course of the
interaction to eventually build a new \emph{in-the-moment} measurement of the
quality of interaction. We finally validate and discuss this metric by comparing
it to manual post-hoc annotations of the video-recordings of the interaction.

%\paragraph{Motivation: Impact on Learning}
%
%Education and learning environments are interesting application fields for
%research on attention assessment, since sustaining children's engagement over
%time is a key variable to ensure effective learning (see for
%instance~\cite{Umbach}), and they indeed motivate this research. In fact, the
%use of robots in education comes with the potential to engage the child in
%meta-cognition through the learning by teaching paradigm, wherein a student
%takes the role of a teacher and experiences stronger educational benefits as a
%result (such as in \cite{Palinscar1984}).
%
%While~\cite{zhao2012learning} recommends the development of a teachable agent
%that is actively, rather than passively, engaged in the learning by teaching
%interaction, the increased learning gain results attained
%in~\cite{okita2006observation} were done so even with the pupil being taught
%having been instructed to not contribute much information. This suggests that it
%may not be necessary, from a learning by teaching point of view, for the
%teachable agent to participate actively in the interaction to incite educational
%benefits. However, from an engagement point of view, this may not necessarily be
%true.
%
%The goal of this research is therefore to \textbf{evaluate the quality of an
%interaction} using the \textbf{visual focus of attention as an engagement
%indicator}. We validate our approach with field experiments conducted in the
%educational context of the CoWriter project~\cite{Hood:2015}, where our
%technique, being based on a regular, low-end RGB camera, avoids the use of
%otherwise typically invasive technologies such as eyetrackers, that would
%negatively impact the natural of the interaction.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Visual Attention Assessment}
\label{sec:vfoa}

\subsection{Related Work}

The relation between one's focus of attention and what he/she is looking at has
long been established~\cite{yarbus1967eye,barber1976perception}, and more
specifically, the existing relationship between gaze and attention during social
interaction, and the related gaze patterns, has been part of classic textbooks
like~\cite{argyle1969social} for decades. As such, there is little doubt that
measuring the direction of gaze is a useful proxy to estimate the (visual) focus
of attention of a social agent, and indeed this is one of the basic tools used
in social psychology.

Estimating attention using gaze is not new to robotics either. A recent survey
by Ruhland~\etal~\cite{ruhland2015review} gives in a broad overview of eye gaze
research in HCI and social robotics. It remains however an active field of
research, as illustrated by several recent
publications~\cite{baxter2014tracking,anzalone,kennedy2015head}.  Performing
such a measure on a robot, in real-time, and in ecologically valid environments
(which rules out bulky or invasive apparatus like eye-trackers, or techniques
requiring fine calibration and/or static interactions) remains a challenge in
HRI.

Looking at techniques that both operate on-line and have been deployed in field
experiments, one finds that most approaches rely on head pose estimation alone
(no eye gaze tracking) and are generally based on depth sensors (RGB-D).
Fanelli~\etal provides an overview of these approaches
in~\cite{fanelli2012real}, and recent examples include~\cite{baxter2014tracking,
anzalone}.

Approaches based on monocular 2D vision have been explored as
well~\cite{peters2010investigating}, with however limited robustness to
occlusions or lightning conditions, and over-reliance on tracking to maintain
real-time performances.  Our work relies on recent advances in template-based
face alignment~\cite{kazemi2014one} that allows fast (in the order of a few
milliseconds) facial feature extraction on 2D images, combined with 3D model
fitting, to obtain a fast, robust and stable 6D head pose estimate, that we
successfully deployed in field experiments involving child-robot interactions.

We derive the field of attention from the head pose: this is supported by
previous work, like~\cite{stiefelhagen2002tracking} that shows that the head
orientation's contribution in overall gaze direction is 68.9\%, which further
translates into a 88.7\% accuracy in estimating the focus of attention from head
pose only in a particular meeting scenario (using eye and head tracking).

While previous preliminary research in HRI seemed on the contrary to indicate
that deriving attentional focus from head pose alone would not be accurate
enough~\cite{kennedy2015head}, we found in our case acceptable levels of
agreement between the robot observations and manual post-hoc annotations, as
detailled hereafter.

\subsection{Head Pose Estimation}

As explained, we derive the visual field of attention from the head pose. Our
technique only involves a single monocular RGB camera used for facial feature
extraction, and a static simplified 3D mesh of a human head.  68 facial features
are extracted using a fast template-based face alignment algorithm by Kazemi and
Sullivan~\cite{kazemi2014one}, as implemented in the open-source {\tt dlib}
library~\cite{dlib09}.  Eight of these features are then matched to their 3D
counterparts (Figure~\ref{head_pose}) and we rely on an iterative $PnP$
algorithm (OpenCV's implementation) to compute the translation and rotation of
the head with respect to the camera frame. With this approach, knowing the
intrinsic parameters of the camera (calibrated camera) is required for an
accurate estimation of the absolute 3D localization of the head.

Besides being fast, the face alignment algorithm has been found to perform well
in terms of robustness, including in a range of difficult situations
(occlusions, large head rotations) as encountered in field experiments.
Figure~\ref{head_pose_real_world} shows a selection of such difficult scenes
with one child.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\linewidth]{head_pose}
    \caption{The 6D head pose is estimated by fitting a 3D model of an
        adult head (left) {\it\sc[FACE REDACTED FOR REVIEW]} onto the detected 2D features of the face (right). We
        rely on an iterative $PnP$ algorithm, using 8 correspondence pairs
        (three are depicted: the sellion -- the nasal depression --, the left
        tragion and the menton). The 3D origin of the head is set at the sellion.}
\label{head_pose}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{head_pose_real_world}
    \caption{Head pose results on images captured during a field experiment.
    Detection of face features (and therefore, estimation of the pose) is robust
    to significant occlusions and face rotations.}
\label{head_pose_real_world}
\end{figure}

\subsection{Field \& Focus of Attention}

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{field_of_attention}
    \caption{\small Screenshot of the real-time attention estimation system. The
    visual field of attention is approximated to a 40$\degree$ cone, spanning
    from the head's sellion. The objects whose 3D pose intersect with this cone are
    considered \emph{in focus}.}

\label{fig:vfoa}
\end{figure}

We model the field of attention as the central region of the field of view.  The
field of view itself is approximated to a cone spanned from the nasal depression
(sellion) of the human face. Different dimensions for the human field of view
can be found in the litterature: Holmqvist~\cite{holmqvist2011eye} models it
with an horizontal aperture of $ \pm 40\degree $ and a vertical aperture of $
\pm 25\degree $, while Spector~\cite{walker1980clinical} for instance suggests
$60\degree$ up, $75\degree$ down, $60\degree$ inwards (towards the nose) and
$95\degree$ outwards.  Previous work on visual perspective taking for social
robotics~\cite{sisbot2011situation} model the field of attention as a cone of
$30\degree$. We retained in this work a slightly wider aperture of $40\degree$.

We then approximate the visual \emph{focus} of attention (VFoA) of the human to
the objects which lie inside this field of attention (Figure~\ref{fig:vfoa}). At
a given time, more than one object can therefore be \emph{in focus}.

Our implementation has two limitations: objects are approximated to points (they
are considered in focus if their \emph{origin} lies in the field of attention),
and we do not check actual visibility: one object could be hidden by another, it
would still be considered as in focus. We did not address these limitations
since our experimental setup (involving relatively small objects with no
occlusions) did not necessitate it. Techniques for more accurate assessment of
the visual perspective of the human peer can be found
in~\cite{sisbot2011situation} for instance.

Within these limitations, computing if object $A(x_A,y_A,z_A)$ is in the field
of attention of the human requires first to transform the coordinates
$A(X_A,Y_A,Z_A)$ into the frame of the face, and then to verify the following
simple inequality (with $fov$ the aperture, and assuming that the main axis
$\vec{x}$ of the field of attention points forward):

\begin{align}
    \sqrt{Y_A^2 + Z_A^2} < tan\left(\frac{fov}{2}\right) \cdot X_A
\label{eq:fov}
\end{align}

Our approach assumes that the pose of the objects of interest are available to
the system: as described in section~\ref{sec:system}, our implementation relies
on the ROS {\sc tf} framework to manage and make available to all software
modules the list of poses of existing objects (represented as {\it frames}), and
dedicated perception modules are in charge of publishing up-to-date informations
regarding the location of the objects of interest (the so-called \emph{situation
assessment}). Due to the nature of the experiment, most of the points of
interest considered for the experimental validation presented hereafter are
static with respect to the robot, thus simplifying the scene perception.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Experimental Validation}
\label{sec:expe}

As presented above, we use the 6D head pose as an approximation of the actual
gaze direction, and we further approximate from here the participant's field of
attention. The assumption that such an approximation of the field of attention
allows to derive the actual focus of attention needs to be validated
experimentally.  Our proposed experiment involves child-robot interactions in
the context of handwriting remediation.  This section details the experimental
procedure and presents our results.

\subsection{Experimental Procedure}

The experiment, part of the \redacted project, involves a robot
which tries to engage a child in handwriting tasks using a \emph{learning by
teaching} paradigm (\ie the child is the teacher, and he/she attempts to improve
the robot's handwriting). A tactile tablet is used as writing support.

Figures~\ref{fig:setup} and~\ref{fig:realSetup} illustrate the experimental
setup: a face-to-face child-robot interaction with an (autonomous) Aldebaran
{\sc nao} robot, in the presence of a facilitator (one of the researchers).

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\columnwidth]{experimental_setup}
    \caption{\small \textbf{Experimental setup}: face-to-face interaction with a {\sc
            nao} robot. The robot writes on the tactile tablet, the child then
            corrects the robot by directly overwriting the words on the tablet
            with a stylus. The facilitator remains next to the child to guide the work. 
            The secondary tablet allows the child to tell the robot what to
            write. The areas of interest -- corresponding to potential target of
            attention -- are circled in red with their name.}
    \label{fig:setup}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\columnwidth]{realSetup}
    \caption{\small Picture of the interaction with one of the children
    {\it\sc[FACE REDACTED FOR REVIEW]}.}
    \label{fig:realSetup}
\end{figure}

The subjects were typically located 50 cm away from the robot with the primary
(writing) tablet in front and the secondary one 30 cm to the left of the first
one.  The facilitator was located about 60 cm to the left of the subject.
Finally, two observers (visible by the child) were located further away from the
interaction field.  Figure~\ref{fig:setup} indicates accordingly the location of
main areas of interest (the two tablets, the robot, the facilitator and the
observers).

The dependent variable is the measurement of the participants' VFoA, assessed in
terms of what the attentional targets of the child are over time. The face of
the child is acquired through a fixed webcam (Logitech {\sf c920}), placed on
the table (see Figure~\ref{fig:setup}), and the attentional targets are then
computed as presented in section~\ref{sec:vfoa}.

%An RGB camera has been used to acquire images of 640x480 pixels at 25 fps. The
%camera was located at 5 cm from the center of nao's feet and its holder was tied
%to a wood base for stability purposes. Thus, the camera's objective was 9 cm
%above and $ 40\degree $ inclination with respect to the surface of the table to
%maximize the detection during the writing time. However, nao's camera could have
%been used, always considering the \textit{tf} frame matrix update according to
%the head orientation.

\subsection{Experimental Procedure}

Six children (ages 5 to 6, 3 boys, 3 girls, none wearing glasses) were enrolled
for this study.  The study took place at school, in an isolated room (the
computer lab). The participants were chosen by the teacher, and would come one
after the other to interact with the robot (duration: $M=19.6$ min, $SD=1.58$).

The interaction is organized in rounds of writing: during a typical round, the
child requests the robot to write something (a single letter, a number, or a
full word), and presents a tactile tablet (equipped with a custom writing
application) to the robot. The robot writes on the tablet by mimicking the
writing gesture (but without actually physically touching the tablet); the child
then pulls back the tablet, corrects the robot's attempt by writing on top of or
next to the robot's writing, and ``sends'' his/her demonstration to the robot by
pressing a small button on the tablet. The robot learns from this demonstration
and tries again. The child continues the turn-taking until they decide to train
the robot on another word. In total, the children performed on average 12.16
($SD=2.61$) rounds of writing (complete details on the rationale and
implementation of this experiment can be found in \redacted).

Once per interaction, the robot interrupts the handwriting task to tell a story
(taking about 2 min), and the turn-based hand-writing task continues afterwards.
The intended purpose of the story-telling episode is to locally trigger a
different set of attention behaviors for the child.

\subsection{System Implementation}

\label{sec:system}

\begin{figure}[h!]
    \centering

\resizebox{0.9\linewidth}{!}{%

\begin{tikzpicture}[
    >=latex,
    node distance=2cm,
    every edge/.style={draw, very thick},
    redarrow/.style={fill=Red, draw=Red},
    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]

%    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};

    \node at (0,0) [cmpt, fill=Green!50] (tf) {tf};
    \node [cmpt, fill=YellowOrange!50, left=of tf] (attention) {attention tracking};
    \node [above=0.6cm of attention, text width=2.7cm,align=center] (input) {\tt /image \\ /camera\_info \\ dlib {\sf faces model} \\ {\sf 3D face template}};
    \node [cmpt, right=of tf] (static) {static pose publishers};
    \node [above=0.6cm of static, text width=2.7cm,align=center] (staticposes) {\sf locations of \\areas of interest};
    \node [cmpt, above=0.6cm of tf] (pose) {robot state publisher};
    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tf] (focus) {focus estimation};
    \node [below=0.6cm of focus] (output) {\tt /frames\_in\_focus};

    \path (staticposes) edge [->] (static);
    \path (static) edge [->] node[label,below] {objects poses} (tf);
    \path (input) edge [->] (attention);
    \path (attention) edge [->] node[label,below] {child face} (tf);
    \path (pose) edge [->] node[label,right] {robot pose} (tf);
    \path (tf) edge [->]  (focus);
    \path (focus) edge [->]  (output);

\end{tikzpicture}
}

    \caption{\small ROS nodes involved in the VFoA estimation (orange nodes were developed for this work).}
    \label{fig:system}
\end{figure}

The experiment was carried out with an Aldebaran {\sc nao} robot, using ROS as a
middleware to build the attention estimation pipeline (Figure~\ref{fig:system}).
Head pose estimation, presented in section~\ref{sec:vfoa}, builds on the {\tt
dlib} and OpenCV libraries; the pose transformations are handled by the ROS {\sc
tf} library. The same {\sc tf} library is used to represent the possible point
of interests as individual frames: an object is considered to be in focus when
its frame lies within the field of attention of the participant
(Figure~\ref{fig:vfoa}).  The implementation is open-source and available at
\redacted.

The implementation of the hand-writing activity itself has been presented by
\redacted.

\subsection{Data Collection \& Analysis}

Successful detections of the head, and, when detected, the attentional targets
of the children as estimated by the robot, were logged during the experiment (in
total, $6\times19.6=117.6$ min of interaction). The only post-processing
consisted in filtering out gaze shifts (short episodes -- below 500ms -- between
two attentional targets).

We video-recorded the interactions, and performed a {\it post-hoc} manual coding
of the focus of attention (24\% double-coded, Cohen's $\kappa=0.91$, high
reliability).  The manual coding forms our attentional \emph{ground-truth}.

To assess the accuracy of the attention estimation by the robot, we computed
over time the overlap between the ground-truth and the robot's estimate and the
inter-rater agreement (Cohen's $\kappa$). The periods where the head was not
detected were \emph{excluded} from the agreement computation: at such times, the
robot explicitly knows that it can not estimate the focus of attention, and as
such, we do not consider that it \emph{wrongly} estimates the focus.

%and the cosine similarity
%matrix.

%The Cronbach's $\alpha$ can be defined as an average of the correlations between 
%the items or values at the same time step \textit{t}. Therefore, results can vary 
%significantly if both sets of values are time shifted. However,
%the cosine similarity becomes a useful measurement in this case as it accounts 
%for temporal delays. Specifically, the ground-truth values acquired though human 
%annotations may not match the measurements of the focuses of attention computed 
%by the robot at the exactly same time step \textit{t} even though the correspondent
%curves may look very similar in shape.

\subsection{Results}

\begin{table*}[ht!]
    \centering
    \caption{\textbf{Attention tracking accuracy}. \emph{Head pose
    tracking} is the percentage of total time of successful detection of the
    head pose; \emph{Agreement} is the percentage of matching time
    between manually annotated focus of attention (ground-truth) and robot's
    computed focus of attention. Total duration: 117.6 min.}

    \begin{tabular}{p{5.5cm}cccccccc}
        \toprule
        {\bf Subject} & 1 & 2 & 3 & 4 & 5 & 6 & {\bf M} & {\it SD} \\
        \midrule
        {\bf Head pose tracking} (\%) & 88.2 & 83.5 & 90.5 & 83.1 & 87.9 & 85.0 & {\bf 86.4} & {\it 3.0} \\ 
        \midrule
        {\bf Agreement} (\%) & 58.9 & 67.1 & 79.2 & 48.3 & 65 & 77.1 & {\bf 65.9} & {\it 11.5}\\
        {\bf Cohen's $\kappa$} & 0.48 & 0.56 & 0.68 & 0.26 & 0.47 & 0.68 & {\bf 0.52} & {\it 0.16}\\
        %{\bf Cosine affinity} & 0.73 & 0.82 & 0.87 & 0.73 & 0.77 & 0.89 & {\bf 0.8} & {\it 0.07} \\
        \bottomrule
    \end{tabular}
    \label{tab:results}
\end{table*}

The main results are reported in Table~\ref{tab:results}
%and Figure~\ref{fig:cosinematrix} (cosine similarity matrix);
Figure~\ref{fig:realExpected} further gives a concrete picture of the
ground-truth \vs computed attentional targets for subject~4 (the subject with
the \emph{least} successful tracking).

%
%\begin{figure}[th!]
%    \centering
%    \includegraphics[width=0.7\columnwidth]{bitmap}
%    \caption{\small Cosine similarity matrix between ground-truth and robot-measured VFoA
%    using a 10 sec. window size. The correspondent numbers are shown
%    in figure \ref{fig:realCaptured}.}
%    \label{fig:cosinematrix}
%\end{figure}

\begin{figure*}[ht!]
    \centering
    \includegraphics[width=\linewidth]{matches-excerpt}
    \caption{\small \textbf{Comparison of computed focus of attention \vs ground
        truth} during a face-to-face child-robot interaction (subject 4
        in table~\ref{tab:results}, 3.5min-long excerpt).
        In blue (top lines), the focus of attention as computed by the robot;
        in orange (bottom lines), the focus of attention as manually annotated
        (ground-truth). The bottom section shows agreement between both (whenever
        the head is detected).}

    \label{fig:realExpected}

    % Maybe we should add the expected VFoA
\end{figure*}

During the whole interaction, the head pose of the children was consistently
tracked, 86\% of the time in average, $SD=3.0$. While this high score is
expected for a face-to-face interaction with a static head-tracking camera
(meaning that the child head would remain in the field of view of the camera
most of the time), this is still comforting in terms of suitability of our
approach for head pose estimation with children in field experiments of this
kind. Expectedly, the primary causes of lost head pose were occlusions with the
hands (similar to the middle-bottom picture in
Figure~\ref{head_pose_real_world}), close proximity with the tablet while
writing, and gaze directed to the facilitator (who was sitting directly on the
left of the child, Figure~\ref{fig:realSetup}).

In terms of attention tracking, Cohen's $\kappa$ values are between 0.47 and
0.68 with one subject resulting in significantly worst tracking, at 0.26. While
the interpretation of Cohen's $\kappa$ is subject to discussion (the number of
the coded values -- in our case 6 -- and the distribution probability of values
-- in our case, values are \emph{not} equiprobable -- are factors impacting
$\kappa$ independently of the level of agreement), the levels of agreement are
\emph{moderate} to \emph{substantial}, with one subject only showing \emph{fair}
agreement~\cite{landis1977measurement}. Further analysis of the videos shows
that the child with the lowest level of agreement was particularly quiet and
would indeed rely more on the eyes to direct his gaze than the other children,
thus leading to a less accurate estimation of his focus of attention.

%\subsubsection{Focus of attention distribution}
%
%Additional information of the head pose can be obtained for the evaluation of
%the VFoA distribution. Figure \ref{fig:heatmap} shows the most salient focuses
%of attention where the peaks are located. These peaks are mainly projected
%between the triangle composed by the robot, the writing tablet and the selection
%tablet using an unsupervised K-means clustering method reveals, considering the
%observers as negligible VFoA.
%
%% Workaround to do subfigure since the packege looks not to work
%\begin{figure}[th!]
%  \centering
%  \begin{tabular}{@{}c@{}}
%    \includegraphics[width=.7\linewidth,height=100pt]{heatmap} \\
%    \small (a)  2D histogram.
%  \end{tabular}
%
%  \vspace{\floatsep}
%
%  \begin{tabular}{@{}c@{}}
%    \includegraphics[width=.7\linewidth,height=100pt]{kmeans} \\
%    \small (b) K-means clustering.
%  \end{tabular}
%
%  \caption{ \small Subject 4 head movements. (head yaw on x-axis, pitch on y-axis).}
%  \label{fig:heatmap}
%\end{figure}

The next section builds upon this technique for real-time estimation of
the focus of attention: by comparing the focus of attention with the set of
attentional targets {\it a priori} expected by the robot, we can estimate to
what extent the user is ``with'' the robot.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{With-me-ness}
\label{sec:withmeness}

\subsection{Concept \& Calculation}

The concept of \emph{with-me-ness} has been introduced in the field of
\emph{Computer Supported Collaborative Learning} (CSCL) by Sharma~\etal
in~\cite{sharma2014me}. Sharma~\etal introduce this concept in an attempt to
answer a recurrent teacher's question: {\it ``how much are the students with
me?''}. They distinguish what they call \emph{perceptual with-me-ness} (the
student follows what the teacher refers to with deictic gestures) from
\emph{conceptual with-me-ness} (the student follows what the teacher refers to
verbally), and they show in an eye-tracking study that \emph{conceptual
with-me-ness} in particular correlates with better learning performance. This
also relates to the concept of gaze cross-recurrence that has been shown to
reflect the quality of the interaction~\cite{jermann2012effects} in
collaborative learning tasks.

Sharma~\etal simply define \emph{conceptual with-me-ness} as the normalized
percentage of time during which the student's gaze overlapped the areas of
teaching slides currently referred to by the teacher.  In order to apply it to
human-robot interactions, we propose to extend this concept, and to define
\emph{conceptual with-me-ness} as the normalized ratio of time that the human
interactant focuses its attention on the attentional target expected by the
robot for the current task (or sub-task).

\begin{algorithm}[h!]
    \centering

    \begin{algorithmic}[1]
    \Procedure{Compute With-me-ness} {}
    \State $d_w, d_e \gets 0$
    \State $t \gets t_{start}$
    \Repeat
    \If{$task(t) \neq$ nil \textbf{and} \par
    \hskip\algorithmicindent$F(task(t)) \neq \varnothing$ \textbf{and} \par
    \hskip\algorithmicindent $f(t) \neq$ nil } \label{algline:skiplosttrack}
    \If{$f(t) \in F(task(t))$}
            \State $d_w \gets d_w + \delta_t$
        \EndIf
        \State $d_e \gets d_e + \delta_t$
    \EndIf
    \State $t \gets t + \delta_t$
    \Until{$t = t_{end}$}

    \State $\mathcal{W}_{[start, end]} \gets \frac{d_w}{d_e}$
    \State \Return $\mathcal{W}_{[start, end]}$
    \EndProcedure

    \end{algorithmic}

    \caption{\textbf{Computation of \emph{with-me-ness}}. $d_w$ stands for the duration
        the human is actually \emph{with} the robot, while $d_e$ stands for the
        total time where the human would be \emph{expected to be with} the robot, $task(t)$
        represents the task performed by the robot at time $t$ (possibly none),
        $F(task)$ represents the (possibly empty) set of expected attentional
        targets associated to task $task$, $f(t)$ represents the actual focus of
        attention of the human measured at time $t$. $\mathcal{W}_{[start, end]}$ represents
        the level of \emph{with-me-ness} from $t_{start}$ to $t_{end}$.}
    \label{alg:with-me-ness}
\end{algorithm}

Algorithm~\ref{alg:with-me-ness} provides a formal way of computing the level of
with-me-ness $\mathcal{W}$ between two time points $[t_{start}, t_{end}]$. A
notable difference with the original definition by Sharma~\etal is that, at a
given time $t$, the task $task(t)$ performed by the robot may elicit more than
one attentional target; thus, at a given time, more than one location can be
regarded as possible \emph{expected} focuses of attention for the human. For
example, a robot which is writing, could typically elicit gazes to its hand as
well as to its head. A human looking at either of these locations would be
considered to be \emph{with} the robot in terms of
interaction\footnote{Considering a probabilistic model of attention expectations
(an attention distribution) would be an interesting extension of this metric.}.
Also notable, we exclude from the computation of $\mathcal{W}$ all of the
periods of time where the user's focus of attention can not be estimated
(typically because the user's face is not visible at those times).

\subsection{Experimental Measure \& Interpretation}

\begin{table}[t]
    \centering
    \caption{Mapping between the interaction phases and the expected
        attentional targets.}

    \begin{tabular}{ll}
        \toprule
        {\bf Phase} & {\bf Expected targets} \\
        \midrule
        Presentation & {\sf robot} \\ 
        Waiting for word to write & {\sf secondary tablet} \\ 
        Writing word & {\sf tablet}, {\sf robot} \\ 
        Waiting for feedback & {\sf tablet}, {\sf secondary tablet} \\ 
        Story telling & {\sf robot} \\ 
        Bye & {\sf robot} \\ 
        \bottomrule
    \end{tabular}
    \label{tab:targets}
\end{table}

Over the course of the experiment presented in section~\ref{sec:expe}, the robot
controller would associate a set of expected attentional targets to the phase of
the interaction (Table~\ref{tab:targets}). For instance, while the robot was
waiting for the child's handwriting demonstration (``\textit{Waiting for
feedback}''), the expected attentional target of the child was the tablet (since
the child was supposed to write there) or the secondary tablet (that displayed a
template of the word, used as a reference by the child). These expected targets
(green lines on Figure~\ref{fig:with-me-ness}) form the robot's attentional {\it
a priori} knowledge and are used to compute the with-me-ness. With-me-ness can
be calculated over the whole interaction or over shorter time windows.
With-me-ness over the whole interaction for the six subjects is reported in
Table~\ref{tab:results-with-me-ness}. The Pearson's correlation with the
ground-truth is $r(4)=0.46$ (significance not computed due to small sample
size). Shorter time windows are interesting for two purposes: to analyse the
level of with-me-ness in relation to specific interaction episodes; to allow a
measurement of with-me-ness by the robot \emph{over the course} of the
interaction (\emph{in-the-moment} measurement) -- in the latter case, one may
typically want to consider a sliding time window.

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{with-me-ness}
    \caption{\small \textbf{With-me-ness}. Evolution of the level of
        \emph{with-me-ness} over the whole $\approx$17min long interaction of
        subject~2. The top chart is similar to Figure~\ref{fig:realExpected}
        with the \emph{expected} attentional targets added in green. The bottom
        diagram represents the instantaneous level of with-me-ness over a
        sliding window of 30 seconds. The blue line is the with-me-ness as estimated
        by the robot, the orange line is the with-me-ness computed from manually
        annotated attentional targets.  Pearson's correlation between both
        series for this subject: $r(973)=0.58, p < .001$.}

    \label{fig:with-me-ness}
\end{figure*}

\begin{table}[h!]
    \centering
    \caption{\textbf{Levels of with-me-ness}. For each subject, the with-me-ness
    level is reported over the whole interaction, either based on the
    ground-truth (\ie \emph{ground-truth with-me-ness}), or based on the focus of
    attention measured by the robot.}

    \begin{tabular}{p{1cm}cccccccc}
        \toprule
        Subject & 1 & 2 & 3 & 4 & 5 & 6 & {\bf M} & {\it SD} \\
        \midrule
        $\mathcal{W}_{g.truth}$ & 79.4 & 81.6  & 90.5 & 87.9 & 90.7 & 80.9 & {\bf 85.2} & {\it 5.1} \\ 
        $\mathcal{W}_{robot}$ & 52.6 & 55.3 & 74.3 & 52.9 & 59.5 & 63.9 & {\bf 59.8} & {\it 8.3} \\
        \bottomrule
    \end{tabular}
    \label{tab:results-with-me-ness}
\end{table}

The with-me-ness plotted at the bottom of Figure~\ref{fig:with-me-ness} is in
fact computed on a sliding window of 30 seconds, and thus gives a picture of
``how well the child is following the robot's expectations'' at that time. As
seen, the with-me-ness computed at run-time by the robot (blue line) is
generally lower than the ground-truth (orange line, based on video-annotations),
and sometimes quite off, such as during episode marked ``A'': during that phase,
one can notice that the attention is mostly directed to undefined target {\sf
Other}, likely a consequence of inaccurate head detection.  This kind of error
(inaccurate head pose estimation) is the main source of discrepancy between the
ground-truth and the attention distribution measured by the robot: ignoring all
the episodes where the child's gaze is measured to be directed to {\sf Other},
we indeed obtain levels of with-me-ness close to the ground-truth (over the six
subjects, $M=87.5, SD=4.6$).

A chart like Figure~\ref{fig:with-me-ness} remains a useful tool to analyse the
interaction, and several observations can be made from it: the green lines
represent how the robot imagine, at a given time, the attention distribution of
the child. They also provide an accurate picture of the overall turn-taking as
viewed by the robot: for instance, the episode ``B'' on
Figure~\ref{fig:with-me-ness} corresponds to one of the ``\emph{Robot writing}''
episodes, surrounded by ``\emph{Waiting for feedback}'' phases like ``C'';
episode ``D'' corresponds to the story telling; etc. In terms of interaction,
the large variance of the duration of these phases reflects the fact that this
child would sometimes take a lot of time to send feedback to the robot, and
sometimes, on the contrary, be very quick.

Looking at the ground-truth focus of attention (orange lines), the first
striking observation is that this child did generally \emph{closely} follow what
the robot was expecting: in that regard, it seems that the child was very much
engaged in the interaction (we discuss in the next section the exact
relationships between with-me-ness and engagement). The only major exception is
the story-telling phase (episode ``D''): the child was seemingly not captivated
by the first half of the story, and their attention was not directed towards the
robot.

Another interesting observation pertains to the facilitator: as one can see,
this child only rarely turned to the facilitator, possibly indicating that the
interaction and the task were meaningful and easy enough for him to follow
alone.

More subtle patterns and events can also be observed: for instance, during the
feedback phases like episode ``C'', we can notice numerous gaze shifts between
the tablet (where the child writes) and the secondary tablet (that showed a
template of the word). The episode ``B'' (robot writing) is also interesting:
the child did not look at the robot, and instead remained focused on the
secondary tablet. This situation is typically useful for the robot to detect as
it may want to adapt its behavior to recover the child's attention.

%A relevant trend that can be observed is the continuous `look for approval'
%situation where the subjects look at the experimenter in order to provide
%feedback about the robot's response before providing their own one (see yellow
%boxes in figure \ref{fig:realExpected}). Such transition tablet-experimenter
%becomes more frequent in several cases, for instance when the experimenter makes
%a suggestion (orange box in figure \ref{fig:realExpected}), in this case, switch
%to a more complicated word.
%
%Once the turn-taking is established, a pattern becomes visible during the
%subject's feedback, when an attempt is performed: Such pattern shows a high
%intermittent gaze frequency from the tablet to the selection. It also reveals
%that the subject is using the demonstration model in the selection tablet to
%provide a better answer to the robot. In fact, subjects decrease the response
%time after each shift as well as the gaze frequency towards the model suggesting
%an assimilation over time. Moreover, when a new demonstration is selected for
%playing, the same behavior is replicated (see yellow boxes in figure
%\ref{fig:realExpected}).

%If we consider the case shown in section \ref{sec:fromFaceTo}, evaluating the
%subject's progression together with the time on task may be an indicator of
%engagement: If the subject answers so quickly with respect to the previous
%attempts with a poorer handwriting may be the case she is not engaged anymore.
%In such hypothetical case an activity switch can be easily implemented in a
%robot architecture. It confirms the findings reported in \cite{anzalone} where
%time response to event is proposed to account quality of interaction. 
%
%However, a high amount of gaze shift (figure \ref{fig:gazeR}) towards the
%demonstration provided can reveal that the challenge is too big or no gaze, the
%exactly the opposite. In this case, the robot response can be adapted to the
%situation. Similar, but least relevant results are obtained using the
%correctness (figure \ref{fig:errorR}) of the child's feedback after each
%attempt. 
%
%\begin{figure}
%  \centering
%  \begin{tabular}{@{}c@{}}
%    \includegraphics[width=.7\linewidth,height=100pt]{gazeR} \\
%    \small (a) Occurrences of a gaze shift.
%  \end{tabular}\label{fig:gazeR}
%
%  \vspace{\floatsep}
%
%  \begin{tabular}{@{}c@{}}
%    \includegraphics[width=.7\linewidth,height=100pt]{errorR} \\
%    \small (b) Word correctness.
%  \end{tabular}\label{fig:errorR}
%
%  \caption{ \small Linear regression of all subject attempts against the gaze
%  shift tablet-model (a) and the correctness as the average euclidean distance
%  of each letter in its corresponding eigenspace with respect to a reference
%  letter (b). The gray region defines the confidence interval.}
%  
%\end{figure}
%
%
%This information can be used to estimate the child's state during the
%interaction. For instance, if we consider that a low level of engagement is
%produced by extremely high or low demanding situations, it is possible to assess
%the engagement level during the interaction based on gaze shift and word
%correctness.

%It is also important to notice that the initiative of the robot into the
%learning process in a turn-taking activity produces an effect in the engagement
%of the human partner. However, experience suggests that a robot that learns
%evoke a positive action in the partner. But the opposite, a robot that is not
%able to progress over time, produces the contrary effect.

%Finally, the subjects set their attention more often to the tablet displaying the generated hand-writing during the robot feedback rather than to the robot itself.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}

\paragraph{Head Pose to Assess Attention: is it Relevant?}

%During the case study, results show that it is possible to distinguish between
%long and short feedback episodes based on the gaze shift and depending on the
%previous attempts performed by the subject. As a result, gaze-shift together
%with the number of attempts becomes a reliable indicator candidate in the
%assessment of the engagement during the activity.
%
%Additionally, both the gaze shift and the correctness of the feedback seem to be
%strictly correlated with the learning gain during the activity as they go in
%decrement after each attempt provided by the subjects. Despite the fact that
%these results are encouraging, there is a need of validation with a greater
%amount of subjects.

We already stated the main limitations of our approach to estimating the focus
of attention: eye gaze information is neglected and we do not perform visibility
check of the in-focus objects (we simply approximate them to their origins,
ignoring possible occlusions).

While the first issue is shared with most of the other vision (2D or 3D) or
motion capture techniques for real-time gaze estimation found in robotics, our
results are positive: we show that relying purely on head pose estimation to
estimate gaze direction leads to real-world measures that are worth being
considered and used. They may not match manual annotations, but they are
definitely a valuable \emph{in-the-moment} input for the robot. For certain
children, we reach levels of accuracy traditionally considered as good.

Interestingly, previous research conducted on ape-human interaction by
Tomasello~\etal~\cite{tomasello2007reliance} shows that apes rely primarily on
the head pose to estimate the focus of attention of a human partner, in to
contrast to human children who primarily rely on eye direction. From this
perspective, our robot mimics attention tracking mechanism comparable to the
ones found in apes.

Our approach relies on a simple, non-intrusive sensor (a RGB camera by the
robot) and an open-source, fast pose estimation algorithm
: we hope that this may contribute to the widespread
adoption of such a technique on a range of robots, including the relatively
common {\sc nao} platform.
%\fixme{test on nao! how fast is it actually? in 320x240, 240ms on nao, 20ms on core i7}

%Gaze information acquired during the interactions is not only useful to detect
%and predict engagement, but also allows the possibility to develop actively the
%concept gaze-aware robot. In other words, make the system able to estimate where
%the subject looks at, but also imitate the same subject's VFoA when the activity
%state allows it.

\paragraph{With-me-ness: yet another metric of engagement?}

Borrowing the neologism from the field of CSCL, we have also introduced in this
article \emph{with-me-ness} as a measure of ``how much the user is \emph{with}
the robot during a task''. This can be acquired over the course of the
interaction, thus providing the robot with a real-time metric for a relatively
high-level social construct, undoubtedly related to engagement.

One may reasonably wonder how different \emph{with-me-ness} is from \emph{joint
attention} on one hand, and from \emph{engagement} on the other hand.
With-me-ness is related to both, with however noteworthy nuances: (Triadic)
\emph{joint attention} is understood as the cognitive realization of a shared
attention to an object, itself building on a shared perception of that object
(\ie joint attention builds on a \emph{perceptual} alignment of two agents).
\emph{Conceptual with-me-ness} as proposed by Sharma~\etal
in~\cite{sharma2014me} is on the contrary \emph{referential}: ``you are
\emph{with} me if you focus on what I refer to, either explicitly or
implicitly''. We understand it here in a slightly broader sense that reflects
the \emph{interaction}: ``you are \emph{with} me if you focus on what is
important for the interactive task at hand.''

On the other hand, with-me-ness is only a precursor of engagement: it does not
say much about the \emph{cognitive} commitment of a user to an interaction. A
user may closely adhere to the injunctions of the robot (or, actually, of the
experimenters), with thus high levels of with-me-ness,  without being
\emph{engaged} in the interaction. This is typically seen in child-robot
interaction: children will attempt to closely follow what they are asked to do
-- which may \emph{look like} they are engaged in the interaction -- while they
merely \emph{obey orders}.

Compared to engagement, one of the strengths of with-me-ness is its specificity:
it is well-defined (the only important variable -- that needs to be reported --
is the mapping {\it phase/expected attentional target}, \ie our
Table~\ref{tab:targets}), we can formalize it, and as such, it is valuable to
assess and compare how users are willing or able to interact with a robot.  We
have hopefully demonstrated in this article that with-me-ness is an operational
\emph{in-the-moment} metric that can be also used as feedback to the robot
controller to build richer, more adaptive interactive behaviors for our robots.

\section{Conclusion}

We have presented how a robot can effectively assess in real-time with a regular
camera, the focus of attention of its interactants, and how we can combine it
with the robot's {\it a priori} knowledge about the interaction to build a
metric of \emph{with-me-ness} over the course of the interaction.

The experimental validation has been conducted with six children in face-to-face
interaction with an autonomous robot, over a total duration of about 2 hours. It
shows that {\it 1)} most of the time, we are able to estimate the head pose of
these children; {\it 2)} based on these head poses only, the instantaneous focus
of attention as computed by the robot does reach a good level of accuracy, with
however one inaccurate outlier out of the six participants; {\it 3)} the robot is
able to compute in real-time a level of \emph{with-me-ness} that correlates
strongly with the ground-truth.

The accuracy of the attention estimation could be improved, first by adding
heuristics to detect and ignore erroneous/inaccurate face detections, second by
implementing pupil tracking on top of head pose estimation. However, our results
show that the accuracy levels that we reach already support the reliable
computation of a metric measuring a high-level social construct, the
\emph{with-me-ness}, which we argued should be used as a well-defined, reliable
precursor of engagement in building adaptive robot behaviors.

\section*{Acknowledgments}

\redacted

\bibliographystyle{abbrv}
\bibliography{sigproc}
\end{document}